---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(reshape2)
library(ggplot2)
library(xts)
library(zoo)
library(tsbox)
library(forecast)
library(tidyverse)
library(ggplot2)
library(psych)
library(fpp2)
```

```{r eval=FALSE}
save.image('/Users/lmp/CNAM/Stats/Mini-memoire/data/all_data.RData')
```

```{r eval=FALSE}
  data_tax_burden <- read_delim('/Users/lmp/CNAM/Stats/Mini-memoire/data/The_Tax_Burden_on_Tobacco__1970-2019.csv', delim = ";")
  str(data_tax_burden)
  dim(data_tax_burden)
  summary(data_tax_burden)
  describe(data_tax_burden)
  head(data_tax_burden)
  unique(data_tax_burden$SubMeasureDesc)
  colnames(data_tax_burden)


  data_1985_2019_cleaned_no_factors <- read.csv("~/CNAM/Stats/Mini-memoire/data/data_1985_2019_cleaned_no_factors.csv")
```

## EDA

## Constitution time series avec les moyennes 

```{r}
    brfss <- data_1985_2019_cleaned_no_factors %>%
      select(-c(X, X_STATE, IYEAR))
    brfss_m <- brfss  %>%
      group_by(IDATE) %>% 
      summarise_all(mean, na.rm = TRUE)
    brfss_dates <- as.Date(brfss_m$IDATE, format = "%Y-%m-%d")

    age_m <- brfss_m[, c("X_AGEG5YR", "X_AGE65YR", "X_AGE_G", "SEX")]
    famille_m <- brfss_m[, c("MARITAL", "PREGNANT", "NUMADULT", "CHILDREN", "X_CHLDCNT")]
    education_m <- brfss_m[, c("EDUCA", "X_EDUCAG", "EMPLOY1")]
    finances_m <- brfss_m[, c("INCOME2", "X_INCOMG", "RENTHOM1")]
    tabac_m <- brfss_m[, c("SMOKE100", "SMOKDAY2", "STOPSMK2", "LASTSMK2", "USENOW3", "X_SMOKER3", "X_RFSMOK3")]
    sante_m <- brfss_m[, c("POORHLTH", "MENTHLTH", "QLMENTL2", "QLSTRES2", "ADDEPEV2", "EMTSUPRT", "LSATISFY")] 
    
    tax_burden <- data_tax_burden %>% select(Year, SubMeasureDesc, Data_Value, Data_Value_Unit)
    tax_m <- tax_burden  %>%
      group_by(SubMeasureDesc, Year) %>% 
      summarise(Data_Value = mean(Data_Value))
    
    tax_rate_m <- tax_m[tax_m$SubMeasureDesc == "Federal and State tax as a Percentage of Retail Price",] 
    tax_dates <- as.Date(ISOdate(tax_rate_m$Year, 1, 1))
    tax_rate_m <- tax_rate_m[,3]
    tax_revenue_m <- tax_m[tax_m$SubMeasureDesc == "Gross Cigarette Tax Revenue",]
    tax_revenue_m <- tax_revenue_m[,3]
    pack_conso_m <- tax_m[tax_m$SubMeasureDesc == "Cigarette Consumption (Pack Sales Per Capita)",]
    pack_conso_m <- pack_conso_m[,3]
    pack_cost_m <- tax_m[tax_m$SubMeasureDesc == "Average Cost per pack",]
    pack_cost_m <- pack_cost_m[,3]
    
    tax_rate <- xts(tax_rate_m, order.by=tax_dates)
    tax_revenue <- xts(tax_revenue_m, order.by=tax_dates)
    pack_conso <- xts(pack_conso_m, order.by=tax_dates)
    pack_cost <- xts(pack_cost_m, order.by=tax_dates)
    
    brfss_xts <- xts(brfss_m, order.by = brfss_dates)
    age <- xts(age_m, order.by=brfss_dates)
    famille <- xts(famille_m, order.by=brfss_dates)
    education <- xts(education_m, order.by=brfss_dates)
    finances <- xts(finances_m, order.by=brfss_dates)
    tabac <- xts(tabac_m, order.by=brfss_dates)
    sante <- xts(sante_m, order.by=brfss_dates)
    sante1 <- sante[, c("POORHLTH", "MENTHLTH", "QLMENTL2", "QLSTRES2")]
    sante2 <- sante[, c("ADDEPEV2", "EMTSUPRT", "LSATISFY")]
```

```{r}
data/stats::lag(finances,-1) - 1
autoplot(diff(finances)/finances[-nrow(finances),] * 100, facets = TRUE)
ts_ggplot(diff(tax_rate)/tax_rate[-nrow(tax_rate),] * 100)
```


- Remplacement des donnees de 2010 par interpolation lineaire

```{r}
original_df <- age
times.init <- as.Date(original_df$DATE, format = "%Y-%m-%d")
new_df <-xts(original_df[,2:4],times.init)

interpol <- na.approx(new_df, xout=seq(min(times.init), max(times.init), "day"))
ts.plot(interpol)

PMS_ts_interpol <- ts_ts(PMS_xts_interpol)
ts_ggplot(PMS_xts_interpol, title="Evolution de la santé mentale des répondants", subtitle="Fumeurs et non fumeurs") +
  theme_tsbox() +
  scale_color_tsbox() # ts_ggplot marche aussi avec xts
```

```{r}
autoplot(famille)
```

## Detection des tendances, saisonnalite et cyclicite & Detection du White Noise

> Trends induce positive correlations in the early lags.
> Seasonality will induce peaks at the seasonal lags.
> Cyclicity induces peaks at the average cycle length.

> Verifier si data est cyclique ou saisonniere :
> When data are either seasonal or cyclic, the ACF will peak around the seasonal lags or at the average cycle length

> White noise :
> p-value greater than 0.05 suggests that the data are not significantly different from white noise
> expectation : each value is close to 0
> 95% of all correlations should line within the blue line

#### Tax burden

```{r}
  autoplot(tabac, facets = TRUE) + # cyclic, no trend
    xlab("Années") + ylab("Pourcentage") +
    ggtitle("Evolution de la taxe par rapport au prix de vente") +
    theme_bw() +
    theme(legend.position="bottom")

  gglagplot(tabac) 
  ggAcf(tax_rate) # pas cyclique
  
  autoplot(tax_revenue, facets = FALSE) # upward trend
  gglagplot(tax_revenue) 
  ggAcf(tax_revenue) # pas cyclique
  
  autoplot(pack_cost, facets = FALSE) # up then downward trend
  gglagplot(pack_cost) 
  ggAcf(pack_cost) # pas cyclique
  
  autoplot(pack_conso, facets = FALSE) # upward trend
  gglagplot(pack_conso)  +
        xlab("Années") + ylab("Pourcentage de taxe par rapport au retail price")
  ggAcf(pack_conso) # pas cyclique
```
```{r}
autoplot(tabac, facets=TRUE) +
        xlab("Années") + ylab("Numéro de la réponse")+
        scale_color_manual(labels = c("Statut matrimonial (1:Marrié, 2:Divorce, 3:Veuf, 4:Séparé, 5:Jamais marrie, 6:En couple non marrie)", 
                                      "Enceinte (1:Oui)", 
                                      "Nombre d'adultes dans le foyer", 
                                      "Nombre d'enfants dans le foyer",
                                      "Nombre d'enfants dans le foyer (0:0, 5: >5 enfant)"), values = c(1, 2, 3, 4, 5, 6,7)) + 
        theme(legend.position="bottom")
```

#### BRFSS

```{r}
  autoplot(age) # upward trend
  gglagplot(age) 
  ggAcf(age) # pas cyclique
  
  autoplot(famille, facets = TRUE) # pas cyclique
  gglagplot(famille) 
  ggAcf(famille) # PREGNANT white noise
  # Ljung-Box Tests
  Box.test(famille[,"PREGNANT"], lag = 24, fitdf = 0, type = "Lj") # confirme WN car p-value = 0.003829
  Box.test(famille[,"MARITAL"], lag = 24, fitdf = 0, type = "Lj") # 2.2e-16
  
  autoplot(education, facets = TRUE) 
  gglagplot(education) 
  ggAcf(education) # pas cyclique
  
  autoplot(finances, facets = TRUE)
  gglagplot(finances) 
  ggAcf(finances) # pas cyclique
  
  autoplot(tabac[,"SMOKE100"])
  gglagplot(tabac[,"SMOKE100"]) 
  ggAcf(tabac) # pas cyclique
  autoplot(tabac, facets = TRUE)
  gglagplot(tabac) 
  ggAcf(tabac) # pas cyclique
  Box.test(tabac[,"SMOKE100"], lag = 2334, fitdf = 0, type = "Lj") # 2.2e-16
  Box.test(tabac[,"SMOKDAY2"], lag = 24, fitdf = 0, type = "Lj") # 2.2e-16
  Box.test(tabac[,"STOPSMOK2"], lag = 24, fitdf = 0, type = "Lj") # 2.2e-16
  Box.test(tabac[,"LASTSMK2"], lag = 24, fitdf = 0, type = "Lj") # 2.2e-16
  Box.test(tabac[,"USENOW3"], lag = 24, fitdf = 0, type = "Lj") # 2.2e-16
  Box.test(tabac[,"X_SMOKER3"], lag = 24, fitdf = 0, type = "Lj") # 2.2e-16
  Box.test(tabac[,"X_RFSMOK3"], lag = 24, fitdf = 0, type = "Lj") # 2.2e-16
  
  autoplot(sante, facets = TRUE)
  gglagplot(sante) 
  ggAcf(sante) # pas cyclique

  finances[index(which.max(finances)),] # premier index de la plus grande valeur
```


```{r}
t <- ts_ts(sante)
ts_ggplot(t, title="Evolution de la santé mentale des répondants", subtitle="Fumeurs et non fumeurs") +
  theme_tsbox() +
  scale_color_tsbox() # ts_ggplot marche aussi avec xts
```

```{r}
autoplot(tabac, series="SMOKE100") +
  autolayer(ma(tabac,5), series="5-MA") +
  xlab("Year") + ylab("GWh") +
  ggtitle("Annual electricity sales: South Australia") +
  scale_colour_manual(values=c("Data"="grey50","5-MA"="red"),
                      breaks=c("Data","5-MA"))
```

```{r}

```









> Tendance : en augmentation
> Seasonality : peut-etre


## Visualisation multivariee

### Correlation

- POORHLTH, MENTHLTH : Nombre de jours
- SMOKE100 : 0 1

#### Caracteristiques

```{r}
# Generate means from eu_percentreturns
colMeans(PMS_ts_interpol)

# Use apply to calculate sample variance from eu_percentreturns
apply(PMS_ts_interpol, MARGIN = 2, FUN = var)

# Use apply to calculate standard deviation from eu_percentreturns
apply(PMS_ts_interpol, MARGIN = 2, FUN = sd)

# Display a histogram of percent returns for each index
par(mfrow = c(2,2))
apply(PMS_ts_interpol, MARGIN = 2, FUN = hist, main = "Tabagisme et sante mentale", xlab = "")

# Display normal quantile plots of percent returns for each index
par(mfrow = c(2,2))
apply(PMS_ts_interpol, MARGIN = 2, FUN = qqnorm, main = "")
qqline(PMS_ts_interpol)
```

#### Relations bivariees

```{r}
poorhlth <- PMS_ts_interpol[,1]
menthlth <- PMS_ts_interpol[,2]
smoke100 <- PMS_ts_interpol[,3]
```

- log returns : quand time trend commune

```{r}
# Make a scatterplot of DAX and FTSE
plot(smoke100, menthlth)
# Make a scatterplot matrix of eu_stocks
pairs(age)

# Convert eu_stocks to log returns
logreturns <- diff(log(PMS_ts_interpol))

# Plot logreturns
plot(PMS_ts_interpol)

# Make a scatterplot matrix of logreturns
pairs(PMS_ts_interpol)
```
#### Covariance et correlation

- covariance positive montre association
  - depend de l'echelle
- correlations est une version standardisee de la covariance
  - ne depend pas de l'echelle
- log returns (si variable petite)
  - valeur tres petite mais correlation tres forte


```{r}
# Use cov() with DAX_logreturns and FTSE_logreturns
cov(poorhlth, menthlth)

# Use cov() with logreturns
cov(PMS_ts_interpol)

# Use cor() with DAX_logreturns and FTSE_logreturns
cor(poorhlth, menthlth)
cov(poorhlth, menthlth) / (sd(poorhlth) * sd (menthlth))

# Use cor() with logreturns
cor(PMS_ts_interpol)
```

#### Autocorrelation estimates at lags

Etudie comment time series etst reliee a son passee
Celles avec une forte autocorrelation sont predictibles

- Lag 1 : 1 jour d'ecart
- Lag 2 : 2 jours d'ecart

ACF autocorrelation function is the autocorrelation function of the time lag
plot de l'acf est skewed a droit car chaque lag est correle avec ses lags recents

```{r}
acf(smoke100, lag.max = 2, plot = FALSE)
acf(poorhlth, lag.max = 2, plot = FALSE)
acf(menthlth, lag.max = 2, plot = FALSE)
acf(PMS_ts_interpol, lag.max = 2, plot = FALSE)

acf(smoke100)
acf(poorhlth)
acf(menthlth)
acf(PMS_ts_interpol)
```
each ACF figure includes a pair of blue, horizontal, dashed lines representing lag-wise 95% confidence intervals centered at zero. These are used for determining the statistical significance of an individual autocorrelation estimate at a given lag versus a null value of zero, i.e., no autocorrelation at 

