---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(reshape2)
library(ggplot2)
library(xts)
library(zoo)
library(tsbox)
```

```{r}
  data_1985_2019 <- read.csv("~/CNAM/Stats/Mini-memoire/data/data_1985_2019_cleaned.csv")
  data <- data_1985_2019
```

## EDA

#### Choix des variables
```{r}
    PMS <- data %>%
      select(IDATE, IYEAR,  SMOKE100 , POORHLTH, MENTHLTH)
    PMS <- drop_na(PMS) # on veut qu'un individu ait repondu a toutes les questions
```

#### Aggregation par date 

```{r}
    PMS_mean <- PMS  %>%
                    filter(IYEAR != 2020) %>%
                    group_by(IDATE) %>% 
                    summarise(POORHLTH = mean(POORHLTH),
                            MENTHLTH = mean(MENTHLTH),
                            SMOKE100 = mean(SMOKE100),
                            DATE = first(IDATE))
```

#### Creation de times series type xts

- Visualisation des donnees originales
```{r}
  PMS_date <- as.Date(PMS_mean$DATE, format = "%Y-%m-%d")
  PMS_cols <- select(PMS_mean, SMOKE100, POORHLTH, MENTHLTH)
  PMS_xts <- xts(PMS_cols, order.by=PMS_date)
  plot(PMS_xts)
```

- Remplacement des donnees de 2010 par interpolation lineaire

```{r}
original_df <-PMS_mean[!startsWith(PMS_mean$IDATE,"2010"),]
times.init <- as.Date(original_df$DATE, format = "%Y-%m-%d")
new_df <-xts(original_df[,2:4],times.init)

PMS_xts_interpol <- na.approx(new_df, xout=seq(min(times.init), max(times.init), "day"))
ts.plot(PMS_xts_interpol)

PMS_ts_interpol <- ts_ts(PMS_xts_interpol)
ts_ggplot(PMS_xts_interpol, title="Evolution de la santé mentale des répondants", subtitle="Fumeurs et non fumeurs") +
  theme_tsbox() +
  scale_color_tsbox() # ts_ggplot marche aussi avec xts
```

> Tendance : en augmentation
> Seasonality : peut-etre


## Visualisation multivariee

### Correlation

- POORHLTH, MENTHLTH : Nombre de jours
- SMOKE100 : 0 1

#### Caracteristiques

```{r}
# Generate means from eu_percentreturns
colMeans(PMS_ts_interpol)

# Use apply to calculate sample variance from eu_percentreturns
apply(PMS_ts_interpol, MARGIN = 2, FUN = var)

# Use apply to calculate standard deviation from eu_percentreturns
apply(PMS_ts_interpol, MARGIN = 2, FUN = sd)

# Display a histogram of percent returns for each index
par(mfrow = c(2,2))
apply(PMS_ts_interpol, MARGIN = 2, FUN = hist, main = "Tabagisme et sante mentale", xlab = "")

# Display normal quantile plots of percent returns for each index
par(mfrow = c(2,2))
apply(PMS_ts_interpol, MARGIN = 2, FUN = qqnorm, main = "")
qqline(PMS_ts_interpol)
```

#### Relations bivariees

```{r}
poorhlth <- PMS_ts_interpol[,1]
menthlth <- PMS_ts_interpol[,2]
smoke100 <- PMS_ts_interpol[,3]
```

- log returns : quand time trend commune

```{r}
# Make a scatterplot of DAX and FTSE
plot(smoke100, menthlth)

# Make a scatterplot matrix of eu_stocks
pairs(PMS_ts_interpol)

# Convert eu_stocks to log returns
logreturns <- diff(log(PMS_ts_interpol))

# Plot logreturns
plot(PMS_ts_interpol)

# Make a scatterplot matrix of logreturns
pairs(PMS_ts_interpol)
```
#### Covariance et correlation

- covariance positive montre association
  - depend de l'echelle
- correlations est une version standardisee de la covariance
  - ne depend pas de l'echelle
- log returns (si variable petite)
  - valeur tres petite mais correlation tres forte


```{r}
# Use cov() with DAX_logreturns and FTSE_logreturns
cov(poorhlth, menthlth)

# Use cov() with logreturns
cov(PMS_ts_interpol)

# Use cor() with DAX_logreturns and FTSE_logreturns
cor(poorhlth, menthlth)
cov(poorhlth, menthlth) / (sd(poorhlth) * sd (menthlth))

# Use cor() with logreturns
cor(PMS_ts_interpol)
```

#### Autocorrelation estimates at lags

Etudie comment time series etst reliee a son passee
Celles avec une forte autocorrelation sont predictibles

- Lag 1 : 1 jour d'ecart
- Lag 2 : 2 jours d'ecart

ACF autocorrelation function is the autocorrelation function of the time lag
plot de l'acf est skewed a droit car chaque lag est correle avec ses lags recents

```{r}
acf(smoke100, lag.max = 2, plot = FALSE)
acf(poorhlth, lag.max = 2, plot = FALSE)
acf(menthlth, lag.max = 2, plot = FALSE)
acf(PMS_ts_interpol, lag.max = 2, plot = FALSE)

acf(smoke100)
acf(poorhlth)
acf(menthlth)
acf(PMS_ts_interpol)
```
each ACF figure includes a pair of blue, horizontal, dashed lines representing lag-wise 95% confidence intervals centered at zero. These are used for determining the statistical significance of an individual autocorrelation estimate at a given lag versus a null value of zero, i.e., no autocorrelation at 

```{r}
# Donnees aggregees par jour
autoplot(tabac[, "SMOKE100"]) + xlab("Annee") + ylab("Pourcentage ayant repondu 'oui'") +
  ggtitle("Avez-vous fume plus de 100 cigarettes dans votre vie ?")
# Donnees aggregees par mois et trimestre
smoke100_month <- apply.monthly(tabac[, "SMOKE100"],FUN=mean)
smoke100_quarter <- apply.quarterly(tabac,FUN=mean)

trend = ma(smoke100_month, order = 4, centre = T)
plot(as.ts(smoke100_month))
lines(trend)
plot(as.ts(trend))

head(ma(smoke100_month, 5), n=10)
tail(ma(smoke100_month, 5), n=10)
ma5 <- ma(smoke100_month, 50)
length(ma5)
ma5[5:15]
ma5[370:385]
plot(ma5)

plot(as.ts(timeserie_beer))
lines(trend_beer)

autoplot(smoke100_month) + xlab("Annee") + ylab("Pourcentage ayant repondu 'oui'") +
  ggtitle("Avez-vous fume plus de 100 cigarettes dans votre vie ?")

autoplot(ts(smoke100_quarter), series="SMOKE100") +
  autolayer(ma(smoke100_quarter, 4), series="5-MA") +
  xlab("Temps") + ylab("Taux de reponse 'Oui' ") +
  ggtitle("Avez-vous fume plus de 100 cigarettes dans votre vie ?")

frequency(smoke100_month)
which.max(smoke100_quarter)
```