---
title: "TS_ARMA_X"
output: html_document
---

## ARIMAX

“Order of integration” is a summary statistic used to describe a unit root process in time series analysis. Specifically, it tells you the minimum number of differences needed to get a stationary series.

ARMAX -> integration = 0
ARIMAX

Difference avec ARMA : “X” and its coefficient β
- β is a coefficient which will be estimated based on the model selection and the data
- "X" exogeneous variable


## ARIMA

#### Autoregression

Chaque observation est regressee sur la precedente
Le modele comprend aussi :
  - RW (slope = 1, non stationnaire, forte persistence, Acf tend vers 0 lentement)

(Today - Mean) = Slope * (Yesterday - Mean) + Noise

Slope (phi) between -1 and 1
- if slope = 0 -> process Y is WN
- if slope ≠ 0 -> Y dependant on both current noise and previous observation + process Yt autocorrelated
- large values of slope phi lead to greater autocorrelation
- negative values of slope result in oscillatory time series
- if slope = 1 and mean = 0 -> process Y is a RW (Today = Yesterday + Noise) which is not stationary
- if slope close to 1 -> AR more persistent, Acf revert to 0 quickly -> values far in the past hav little impact

> Independant variable : sante mentale
> Dependant variable : tabac

#### Autoregression 2 : y a t il de la persistence ? 
 > Nombre de fumeurs baisse-t-il de maniere persistente ?

Persistence is defined by a high correlation between an observation and its lag, while anti-persistence is defined by a large amount of variation between an observation and its lag.

```{r}
# Copy and paste the slope (ar1) estimate
0.3140

# Copy and paste the slope mean (intercept) estimate
0.4807

# Copy and paste the innovation variance (sigma^2) estimate
0.002064

# Fit the AR model to AirPassengers
AR <- arima(smoke100, order = c(1,0,0))
print(AR)

# Run the following commands to plot the series and fitted values
ts.plot(smoke100)
AR_fitted <- smoke100 - residuals(AR)
points(AR_fitted, type = "l", col = 2, lty = 2)
```
#### Prediction a partir du AR

```{r}
# Fit an AR model to Nile
AR_fit <- arima(smoke100, order  = c(1, 0, 0))
print(AR_fit)

# Use predict() to make a 1-step forecast
predict_AR <- predict(AR_fit)

# Obtain the 1-step forecast using $pred[1]
predict_AR$pred[1]

# Use predict to make 1-step through 10-step forecasts
predict(AR_fit, n.ahead = 10)

# Run to plot the Nile series plus the forecast and 95% prediction intervals
ts.plot(smoke100, xlim = c(1996, 2050))
AR_forecast <- predict(AR_fit, n.ahead = 10000)$pred 
AR_forecast_se <- predict(AR_fit, n.ahead = 10000)$se
points(AR_forecast, type = "l", col = 2)
points(AR_forecast - 2*AR_forecast_se, type = "l", col = 2, lty = 2)
points(AR_forecast + 2*AR_forecast_se, type = "l", col = 2, lty = 2)

```
#### Auto modele

```{r}
# Fit an automatic ARIMA model to the austa series
fit <- auto.arima(smoke100)

# Check that the residuals look like white noise
checkresiduals(fit)
residualsok <- TRUE

# Summarize the model
summary(fit)

# Find the AICc value and the number of differences used


# Plot forecasts of fit
fit %>% forecast(h = 10) %>% autoplot()
```



## Stationnary or not 
mean, variance, covariance constant over time

