---
title: "Exercise and Tobacco"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)

load("brfss2013.RData")
```

# exerany2: Exercise In Past 30 Days
# exract11: Type Of Physical Activity
# exeroft1: How Many Times Walking, Running, Jogging, Or Swimming
# exerhmm1: Minutes Or Hours Walking, Running, Jogging, Or Swimming
# exract21: Other Type Of Physical Activity Giving Most Exercise During Past Month
# exeroft2: How Many Times Walking, Running, Jogging, Or Swimming
# exerhmm2: Minutes Or Hours Walking, Running, Jogging, Or Swimming
# strength: How Many Times Did You Do Physical Activities Or Exercises To Strengthen Your Mu
# _totinda: Leisure Time Physical Activity Calculated Variable
# metvl11_: Activity Met Value For First Activity
# metvl21_: Activity Met Value For Second Activity
# maxvo2_: Estimated Age-Gender Specific Maximum Oxygen Consumption
# fc60_: Estimated Functional Capacity
# actin11_: Estimated Activity Intensity For First Activity
# actin21_: Estimated Activity Intensity For Second Activity
# padur1_: Minutes Of First Activity
# padur2_: Minutes Of Second Activity
# pafreq1_: Physical Activity Frequency Per Week For First Activity
# pafreq2_: Physical Activity Frequency Per Week For Second Activity
# _minac11: Minutes Of Physical Activity Per Week For First Activity
# _minac21: Minutes Of Physical Activity Per Week For Second Activity
# strfreq_: Strength Activity Frequency Per Week
# pamiss1_: Missing Physical Activity Data
# pamin11_: Minutes Of Physical Activity Per Week For First Activity
# pamin21_: Minutes Of Physical Activity Per Week For Second Activity
# pa1min_: Minutes Of Total Physical Activity Per Week
# pavig11_: Minutes Of Vigorous Physical Activity Per Week For First Activity
# pavig21_: Minutes Of Vigorousphysical Activity Per Week For Second Activity
# pa1vigm_: Minutes Of Total Vigorous Physical Activity Per Week
# _pacat1: Physical Activity Categories
# _paindx1: Physical Activity Index
# _pa150r2: 150 Minute Physical Activity Calculated Variable
# _pa300r2: 300 Minute Physical Activity Calculated Variable
# _pa30021: 300 Minute Physical Activity 2-Level Calculated Variable
# _pastrng: Muscle Strengthening Recommendation
# _parec1: Aerobic And Strengthening Guideline
# _pastae1: Aerobic And Strengthening (2-Level)

```{r}
exercise <- subset(brfss2013, select=c("exerany2", "exract11", "exeroft1", 
                                       "exerhmm1", "exract21", "exeroft2", 
                                       "exerhmm2", "strength", "X_totinda",
                                       "metvl11_", "metvl21_", "maxvo2_",
                                       "fc60_", "actin11_", "actin21_",
                                       "padur1_", "padur2_", "pafreq1_",
                                       "pafreq2_", "X_minac11", "X_minac21",
                                       "strfreq_", "pamiss1_", "pamin11_",
                                       "pamin21_", "pa1min_", "pavig11_",
                                       "strfreq_", "pamiss1_", "pamin11_",
                                       "pavig21_", "pa1vigm_", "X_pacat1",
                                       "X_paindx1", "X_pa150r2", "X_pa300r2",
                                       "X_pa30021", "X_pastrng", "X_parec1",
                                       "X_pastae1"))
```

