---
title: "Health and Tobacco"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)

load("brfss2013.RData")
```

# bphigh4: Ever Told Blood Pressure High
# bpmeds: Currently Taking Blood Pressure Medication
# _rfhype5: High Blood Pressure Calculated Variable

```{r}
hypertension <- subset(brfss2013, select=c("bphigh4", "bpmeds", "X_rfhype5"))
```

# cvdinfr4: Ever Diagnosed With Heart Attack
# cvdcrhd4: Ever Diagnosed With Angina Or Coronary Heart Disease
# cvdstrk3: Ever Diagnosed With A Stroke
# asthma3: Ever Told Had Asthma
# asthnow: Still Have Asthma
# _ltasth1: Lifetime Asthma Calculated Variable
# _casthm1: Current Asthma Calculated Variable
# _asthms1: Computed Asthma Status
# chcocncr: (Ever Told) You Had Any Other Types Of Cancer?
# chccopd1: (Ever Told) You Have (Copd) Chronic Obstructive Pulmonary Disease, Emphysema Or
# havarth3: Told Have Arthritis
# addepev2: Ever Told You Had A Depressive Disorder

```{r}
chronic_health_disorder <- subset(brfss2013, select=c("cvdinfr4", "cvdcrhd4", "cvdstrk3", 
                                                      "asthma3", "asthnow", "chcocncr", 
                                                      "chccopd1", "havarth3", "addepev2",
                                                      "X_ltasth1", "X_casthm1", "X_asthms1"))
```

# physhlth: Number Of Days Physical Health Not Good

```{r}
physical_healthy_days <- subset(brfss2013, select=c("physhlth"))

```

# genhlth: General Health

```{r}
general_health <- subset(brfss2013, select=c("genhlth"))
```

# harehab1: Outpatient Rehab After Heart Attack Hosp Stay
# strehab1: Outpatient Rehab After Hosp Stay For Stroke
# cvdasprn: Take Aspirin Daily Or Every Other Day
# aspunsaf: Health Makes Taking Aspirin Unsafe
# rlivpain: Take Aspirin To Relieve Pain
# rduchart: Take Aspirin To Reduce Chance Of Heart Attack
# rducstrk: Take Aspirin To Reduce Chance Of Stroke

```{r}
cardiovascular <- subset(brfss2013, select=c("harehab1", "strehab1", "cvdasprn", 
                                             "aspunsaf", "rlivpain", "rduchart", 
                                             "rducstrk"))
```


