---
title: "Health Care Access and Tobacco"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggplot2)

load("brfss2013.RData")
```

# hlthpln1: Have Any Health Care Coverage
# persdoc2: Multiple Health Care Professionals
# medcost: Could Not See Dr. Because Of Cost
# checkup1: Length Of Time Since Last Routine Checkup
# medicare: Do You Have Medicare?
# hlthcvrg: Health Insurance Coverage
# delaymed: Delayed Getting Medical Care
# dlyother: Delayed Getting Medical Care Other Response
# nocov121: Without Health Care Coverage Past 12 Months
# lstcovrg: Time Since Last Had Health Care Coverage
# drvisits: Doctor Visits Past 12 Months
# medscost: Could Not Get Medicine Due To Cost
# carercvd: Satisfied With Care Received
# medbills: Currently Have Medical Bills
# _hcvu651: Respondents Aged 18-64 With Health Care Coverage

```{r}
health_care_access <- subset(brfss2013, select=c("hlthpln1", "persdoc2", "medcost",
                                                 "checkup1", "medicare", "hlthcvrg",  
                                                 "delaymed", "dlyother", "nocov121",
                                                 "lstcovrg", "drvisits", "medscost",
                                                 "carercvd", "medbills", "X_hcvu651"))
```

```{r}
str(health_care_access)
```























